---
title: "Music APIs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is based on a custom package I'm developing in parallel: *songwriteR*.  You can install it like this:
```{r, eval=FALSE}
devtools::install_github('ajrominger/songwriteR')
```

## *Spotify*

First and foremost we should note that *Spotify* no longer supports location queries with its API.  Doing a little digging about how that could be done I found [this](https://stackoverflow.com/questions/11367264/solutions-for-finding-website-visitors-geolocation) and [this](http://freegeoip.net/?q=192.12.12.251).  I'll come back to that if we decide we'd like to get spatial data for *Spotify*.

We can still use the *Spotify* API as is to access a metric of popularity, connections between songs, and the sonic "mood" of a song. 

Before using the API we need to setup a dev account with *Spotify*. To do so follow the instructions [here](https://developer.spotify.com/web-api/tutorial/), starting at "Setting Up Your Account" (follow the re-direct link).

For using the *spotifyr* package we can save the account info to our `R` system:
```{r, eval=FALSE}
Sys.setenv(SPOTIFY_CLIENT_ID = [YOUR_CLIENT_ID])
Sys.setenv(SPOTIFY_CLIENT_SECRET = [YOUR_CLIENT_SECRET])
```

```{r, echo=FALSE}
Sys.setenv(SPOTIFY_CLIENT_ID = '29779f58a9be47a48f0ee6a672d229b1')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '517774bfdcdc43688255cf00a5e65f3e')
```

Now we're ready to go.  Let's start by just looking at one artist:

```{r}
library(spotifyr)
# note that a given name could match multiple artist names, so we need to deal with that
# here i just take the first entry
willie <- get_artists('willie nelson')$artist_uri[1]
```

## Genius

[*Genius*](https://genius.com/) is a music aggregator and we can use it to access lyrics, borrowing *RCharlie's* code:

> Similar to with Spotify, I first used the `search` API call to get the `artist_id`. Go [here](https://genius.com/signup_or_login) to set up a dev account to get an API token.
>
>```{r, eval=FALSE}
token <- 'xxxxxxxxxxxxxxxxxxxx'

genius_get_artists <- function(artist_name, n_results = 10) {
    baseURL <- 'https://api.genius.com/search?q=' 
    requestURL <- paste0(baseURL, gsub(' ', '%20', artist_name),
                         '&per_page=', n_results,
                         '&access_token=', token)
    
    res <- GET(requestURL) %>% content %>% .$response %>% .$hits
    
    map_df(1:length(res), function(x) {
        tmp <- res[[x]]$result$primary_artist
        list(
            artist_id = tmp$id,
            artist_name = tmp$name
        )
    }) %>% unique
}

genius_artists <- genius_get_artists('radiohead')
```
>
>Next, I looped through the contents of the songs endpoint (the limit is 50 per page), pulling down each result (a list containing the url of the tracks’ lyrics) until the next_page parameter was null.
>
>```{r, eval=FALSE}
baseURL <- 'https://api.genius.com/artists/' 
requestURL <- paste0(baseURL, genius_artists$artist_id[1], '/songs')

track_lyric_urls <- list()
i <- 1
while (i > 0) {
    tmp <- GET(requestURL, query = list(access_token = token, per_page = 50, page = i)) %>% content %>% .$response
    track_lyric_urls <- c(track_lyric_urls, tmp$songs)
    if (!is.null(tmp$next_page)) {
        i <- tmp$next_page
    } else {
        break
    }
}
```
>
>From here, I used *rvest* to scrape the "lyrics" elements from the urls provided above.
>
>```{r, eval=FALSE}
library(rvest)

lyric_scraper <- function(url) {
    read_html(url) %>% 
        html_node('lyrics') %>% 
        html_text
}

genius_df <- map_df(1:length(track_lyric_urls), function(x) {
    # add in error handling
    lyrics <- try(lyric_scraper(track_lyric_urls[[x]]$url))
    if (class(lyrics) != 'try-error') {
        # strip out non-lyric text and extra spaces
        lyrics <- str_replace_all(lyrics, '\\[(Verse [[:digit:]]|Pre-Chorus [[:digit:]]|Hook [[:digit:]]|Chorus|Outro|Verse|Refrain|Hook|Bridge|Intro|Instrumental)\\]|[[:digit:]]|[\\.!?\\(\\)\\[\\],]', '')
        lyrics <- str_replace_all(lyrics, '\\n', ' ')
        lyrics <- str_replace_all(lyrics, '([A-Z])', ' \\1')
        lyrics <- str_replace_all(lyrics, ' {2,}', ' ')
        lyrics <- tolower(str_trim(lyrics))
    } else {
        lyrics <- NA
    }
    
    tots <- list(
        track_name = track_lyric_urls[[x]]$title,
        lyrics = lyrics
    )
    
    return(tots)
})
```
